A-maze: Easier measurement of incremental processing

### 2021-04-21 Papers/UCI_2021/uci_2021.tex
updates

\documentclass[12pt, xcolor=beamer,table,usenames,dvipsnames, ignorenonframetext, ngerman]{beamer}
\usetheme{Frankfurt}
\usecolortheme{dove}
\usepackage{appendixnumberbeamer}
%\setbeamersize{text margin left=20pt,text margin right=20pt,}

@@

\begin{frame}{Road map} %Acknowledge collaborators 
	%TODO: find a picture of Richard!!
	Methods talk 
	- briefly summarize RT methods, introduce Maze task
	- talk through study 1 (w/ Roger and Richard) where we created and tested this variant method
	- talk through study 2 where I tweaked the method and did another study
	- have a little historical revisionism -- there were some boring implementational tweaks, so what I'll present is a little different that what was used in the first study
	\begin{tikzpicture}[remember picture,overlay]
	\node[xshift=-1.8cm,yshift=-1cm] at (current page.east) {\includegraphics[clip, trim=.1cm 0cm .2cm 0cm,width=.2\textwidth]{../Images/roger.jpg}};
	\end{tikzpicture}

	\begin{tikzpicture}[remember picture,overlay]
	\node[xshift=2.5cm,yshift=-1.5cm] at (current page.north west) {\includegraphics[width=.3\textwidth]{../Images/cpl_header.png}};
	\end{tikzpicture}

@@

Examples of increased difficulty
\begin{itemize}\pause
	\item Constructions not in the grammar \pause
	\item Lexical items that are harder to retrieve \pause
	\item Words that force reparsing or reanalysis
\end{itemize}
\pause
\medskip

@@

\begin{tikzpicture} 
\draw (-.5,-.5) rectangle (3.2,.5);
\onslide<4>{\node at (0,0) {The};}
\onslide<4>{\node at (1.7,0) {- - - - - - - - - - -};}
\onslide<5>{\node at (0,0) {- - - };}
\onslide<5>{\node at (.7,0) {cat};}
\onslide<5>{\node at (2,0) {- - - - - - - - };}
\onslide<6>{\node at (0.3,0) {- - - - - - };}
\onslide<6>{\node at (1.6,0) {drank};}
\onslide<6>{\node at (2.6,0) {- - - -};}
\onslide<7->{\node at (.9,0) {- - - - - - - - - - -};}
\onslide<7->{\node at (2.7,0) {milk.};}
\end{tikzpicture}
\end{center}
\onslide<8->{
			\begin{itemize}
	\item Lots of spillover
	\item Messy data 
\end{itemize}
}
\end{column}
\end{columns}
	\bigskip
\onslide<8->{Different methods have different trade-offs}
\end{frame}

@@

	\pause
	Wrote an Ibex module \pause

	\begin{center}\includegraphics[width=.7\textwidth]{../Images/screenshot.png} \end{center}\pause

@@

\begin{frame}{Meanwhile in Natural Language Processing}
\pause
Language models (LMs)
\begin{itemize}
	\item Trained on large corpora to predict the next word
	\item Given a partial sentence, return probabilities of the next word
\end{itemize}
\pause
Surprisal: negative log probability
\begin{itemize}
	\item 2 bits of surprisal = 1/4
	\item 10 bits of surprisal $\approx$ 1/1000 
	\item +1 surprisal = half as likely
\end{itemize}
\pause
\begin{tikzpicture}
\node (A) at (0,0) {"The dog chased ..."};
\node  (B) at (4,0) [draw, minimum width=1.5cm, minimum height=1.5cm, fill=black] {\textcolor{white}{LM}} ;
\draw [->, thick] (A) -- (B); 
\node[align=left] (C) at (7,0) {"the" 1.2 bits \\"a" 3.1 bits\\"cats" 8 bits\\...\\"wish" 25 bits\\...};
\draw [->, thick] (B) -- (C); 
\end{tikzpicture}
\end{frame}

@@

\begin{frame}{Does it work?} 
	%\vskip -2em
	\small
	From ``Maze Made Easy'' (Boyce et al 2020)
	\medskip

@@

\begin{frame}{Does it work?} 
	\pause
	%\vskip -2em
	{\large Yes, at least well enough.} \pause
	\begin{itemize}
		\item Sometimes generates plausible distractors. \pause
		\item Sloggett et al 2020 also found A-maze results comparable with G-maze
	\end{itemize} 
\end{frame}

@@

	Incidentally, also ``solves'' problem of bad distractors!

@@

\begin{frame}{Surprisal Effects}
	\includegraphics[width=.9\textwidth]{../Images/amaze_gam.pdf}	
\end{frame}

\begin{frame}{SPR comparison}

@@

\end{frame}
\begin{frame}{Surprisal Effects}
	Takeaways: \pause
	\begin{itemize}
		\item Minimal frequency effects (consistent with Shain 2019) \pause
		\item Large effects of Length, Surprisal \pause
		\item Very little spillover \pause
	\end{itemize}
	%	
	Model comparison: GRNN is best, but TXL complementary
\end{frame}

@@

\begin{frame}{Consider A-maze!}
Easy to use! \pause
\begin{itemize}
	\item Runs on command line \pause
	\item Match distractors across minimal pair sentences \pause
	\item Customize surprisal thresholds, vocabulary lists \pause
	\item Can output pre-formatted for Ibex
\end{itemize}
\medskip
\end{frame}

\begin{frame}{Consider A-maze!} \pause
Adapt A-maze to your projects:
\begin{itemize}
	\item Written in Python 3 
	\item Interface with other language models 
	\item Add more frequency sources 
	\item Extend to non-English languages
\end{itemize}

@@

\item Paper: psyarxiv.com/b7nqd/
\end{itemize}
\end{frame}

@@

\begin{frame}{Why such large effects?}
	\pause
	Bayesian Reader (Norris 2006): Look at words long enough to ID with some threshold of certainty \pause

	Possible mechanisms for difference: \pause
	\begin{itemize}
		\item Higher threshold \pause 
		\item Fewer available resources for processing \pause
		\item Presence of second word 
	\end{itemize}

>>>>

\documentclass[ 12pt, xcolor=beamer,table,usenames,dvipsnames, ignorenonframetext, ngerman]{beamer}
\usetheme{Frankfurt}
\usecolortheme{dove}
\usepackage{appendixnumberbeamer}
%\setbeamersize{text margin left=20pt,text margin right=20pt,}

@@

\begin{frame}{Plan} %Acknowledge collaborators 
	\pause
%Methods \& experiments to test them
\bigskip
\begin{itemize}
	\item Maze task
	\item A-maze
	\item Study 1: methods comparison
	\item Variant of A-maze
	\item Study 2: test on Natural Stories Corpus
\end{itemize}\pause 
	\begin{tikzpicture}[remember picture,overlay]
	\node[xshift=-2cm,yshift=0cm] at (current page.east) {\includegraphics[clip, trim=.1cm 0cm .2cm 0cm,width=.2\textwidth]{../Images/roger.jpg}};
	\end{tikzpicture}
		\begin{tikzpicture}[remember picture,overlay]
	\node[xshift=-2cm,yshift=2cm] at (current page.south east) {\includegraphics[clip, trim=.1cm 0cm .2cm 0cm,width=.2\textwidth]{../Images/richard.jpg}};
	\end{tikzpicture}
	\begin{tikzpicture}[remember picture,overlay]
	\node[xshift=-2cm,yshift=-2cm] at (current page.north east) {\includegraphics[width=.3\textwidth]{../Images/cpl_header.png}};
	\end{tikzpicture}

@@

Examples of increased difficulty
\begin{itemize}
	\item Constructions not in the grammar 
	\item Lexical items that are harder to retrieve 
	\item Words that force reparsing or reanalysis
\end{itemize}
\pause
\medskip

@@

\begin{tikzpicture} 
\draw (-.5,-.5) rectangle (3.2,.5);
\onslide<4>{\node at (0,0) {The};}
\onslide<4>{\node at (1.7,0) {- - - - - - - - - - -};}
\onslide<5>{\node at (0,0) {- - - };}
\onslide<5>{\node at (.7,0) {cat};}
\onslide<5>{\node at (2,0) {- - - - - - - - };}
\onslide<6>{\node at (0.3,0) {- - - - - - };}
\onslide<6>{\node at (1.6,0) {drank};}
\onslide<6>{\node at (2.6,0) {- - - -};}
\onslide<7->{\node at (.9,0) {- - - - - - - - - - -};}
\onslide<7->{\node at (2.7,0) {milk.};}
\end{tikzpicture}
\end{center}
\onslide<8->{
			\begin{itemize}
	\item Lots of spillover
	\item Messy data 
\end{itemize}
}
\end{column}
\end{columns}
	\bigskip
\onslide<9->{Different methods have different trade-offs}
\end{frame}

@@

	\pause
	\begin{columns}
		\begin{column}{0.5\textwidth}
				Wrote an Ibex module
		\end{column}
		\begin{column}{0.5\textwidth} 
			\begin{center}\includegraphics[width=\textwidth]{../Images/screenshot.png} \end{center}\pause
		\end{column}
	\end{columns}

@@

\begin{frame}{Meanwhile in Natural Language Processing}
\pause
Language models (LMs)
\begin{itemize}
	\item Trained on large corpora to predict the next word
	\item Given a partial sentence, return probabilities of the next word
\end{itemize}
Surprisal: negative log probability
\begin{itemize}
	\item 2 bits of surprisal = 1/4
	\item 10 bits of surprisal $\approx$ 1/1000 
	\item +1 surprisal = half as likely
\end{itemize}
\begin{tikzpicture}
\node (A) at (0,0) {"The dog chased ..."};
\node  (B) at (4,0) [draw, minimum width=1.5cm, minimum height=1.5cm, fill=black] {\textcolor{white}{LM}} ;
\draw [->, thick] (A) -- (B); 
\node[align=left] (C) at (7,0) {"the" 1.2 bits \\"a" 3.1 bits\\"cats" 8 bits\\...\\"wish" 25 bits\\...};
\draw [->, thick] (B) -- (C); 
\end{tikzpicture}
\end{frame}

@@

\begin{frame}{Does it work?} 
	%\vskip -2em
	\medskip

@@

\begin{frame}{Does it work?} 
	\pause
	%\vskip -2em
	{\large Yes, at least well enough.} \pause
	\begin{itemize}
		\item Caveat: Sometimes generates plausible distractors. \pause
		\item Sloggett et al (2020) also found A-maze results comparable with G-maze
	\end{itemize} 
\end{frame}

@@

%	Incidentally, also ``solves'' problem of bad distractors!

@@

\begin{frame}{Surprisal Effects}
	\includegraphics[width=\textwidth]{../Images/amaze_gam.pdf}	
\end{frame}

\begin{frame}{SPR comparison}
	Using SPR data collected by Futrell et al. (2017)

@@

\end{frame}
\begin{frame}{Surprisal Effects}
	Takeaways: 
	\begin{itemize}
		\item Minimal frequency effects (consistent with Shain 2019) 
		\item Large effects of Length, Surprisal
		\item Very little spillover 
	\end{itemize}
	%	
%	Model comparison: GRNN is best, but TXL complementary
\end{frame}

@@

\begin{frame}{Consider A-maze!}
Easy to use! \pause
\begin{itemize}
	\item Runs on command line \pause
	\item Match distractors across minimal pair sentences \pause
	\item Customize surprisal thresholds, vocabulary lists \pause
	\item Can output pre-formatted for Ibex\pause
\end{itemize}

Adapt A-maze to your projects:
\begin{itemize}
	\item Written in Python 3 
	\item Interface with other language models 
	\item Add more frequency sources 
	\item Extend to non-English languages
\end{itemize}
\medskip
\end{frame}

@@

\item Exp 1 Paper: psyarxiv.com/b7nqd/
\end{itemize}
\end{frame}

@@

\begin{frame}{Why such large effects?}
	Bayesian Reader (Norris 2006): Look at words long enough to ID with some threshold of certainty

	Possible mechanisms for difference: 
	\begin{itemize}
		\item Higher threshold 
		\item Fewer available resources for processing
		\item Presence of second word 
	\end{itemize}
